---
title: "MA 347 Final Project"
output: html_document
---

```{r}
library(magrittr)
library(dplyr)
library(rpart)
library(rpart.plot)
library(caret)
```

# Transforming the data


```{r}
germancredit <- read.csv("~/GermanCredit.csv")
germancredit$purpose_credit <- rep(0, dim(germancredit)[1]) #new column for all purpose of credit variables

idx <- which(germancredit$NEW_CAR == 1)
germancredit$purpose_credit[idx] <- "New Car"

idx <- which(germancredit$USED_CAR == 1)
germancredit$purpose_credit[idx] <- "Used Car"

idx <- which(germancredit$FURNITURE == 1)
germancredit$purpose_credit[idx] <- "Furniture"

idx <- which(germancredit$RADIO.TV == 1)
germancredit$purpose_credit[idx] <- "Radio/TV"

idx <- which(germancredit$EDUCATION == 1)
germancredit$purpose_credit[idx] <- "Education"

idx <- which(germancredit$RETRAINING == 1)
germancredit$purpose_credit[idx] <- "Retraining"

idx <- which(germancredit$purpose_credit == 0)
germancredit$purpose_credit[idx] <- "Other"

germancredit$purpose_credit <- as.factor(germancredit$purpose_credit)
```


```{r}
germancredit$marital_status <- rep(0, dim(germancredit)[1]) # new column for marital status variables

idx <- which(germancredit$MALE_DIV == 1)
germancredit$marital_status[idx] <- "Male and Divorced"

idx <- which(germancredit$MALE_SINGLE == 1)
germancredit$marital_status[idx] <- "Male and Single"

idx <- which(germancredit$MALE_MAR_or_WID == 1)
germancredit$marital_status[idx] <- "Male and Married/Widowed"

idx <- which(germancredit$marital_status == 0)
germancredit$marital_status[idx] <- "Other"

germancredit$marital_status <- as.factor(germancredit$marital_status)
```

```{r}
germancredit$ownership_status <- rep(0, dim(germancredit)[1]) # new column column for owning and renting

idx <- which(germancredit$OWN_RES == 1)
germancredit$ownership_status[idx] <- "Own Residence"

idx <- which(germancredit$RENT == 1)
germancredit$ownership_status[idx] <- "Rent Residence"


idx <- which(germancredit$ownership_status == 0)
germancredit$ownership_status[idx] <- "Other"

germancredit$ownership_status <- as.factor(germancredit$ownership_status)
```

```{r}
germancredit$realestate <- rep(0, dim(germancredit)[1]) # new column real estate and no real estate

idx <- which(germancredit$REAL_ESTATE == 1)
germancredit$realestate[idx] <- "Yes"

idx <- which(germancredit$PROP_UNKN_NONE == 1)
germancredit$realestate[idx] <- "No/unknown"


idx <- which(germancredit$realestate == 0)
germancredit$realestate[idx] <- "Other"

germancredit$realestate <- as.factor(germancredit$realestate)
```

# Splitting the dataset

```{r}
set.seed(2)
train.index <- sample(c(1:dim(germancredit)[1]), dim(germancredit)[1]*0.7)  
train.df <- germancredit[train.index, ]
valid.df <- germancredit[-train.index, ]
```

# CART Model
Making the classification tree
```{r}
set.seed(1)
class_tr <- rpart(RESPONSE ~ CHK_ACCT + HISTORY + purpose_credit + SAV_ACCT + EMPLOYMENT + realestate + FOREIGN, 
                  data = train.df, 
                  minsplit = 1,
                    method = "class") 
```

finding the optimal cp
```{r}
lowest_cp<-class_tr$cptable[which.min(class_tr$cptable[,"xerror"]),"CP"]
lowest_cp
```

Pruning the tree
```{r}
pruned_tr <- prune(class_tr, cp = 0.01530612 )
```


Checking classification rate of training data
```{r}
treepred <- predict(pruned_tr, newdata = train.df, type = "class")
confusionMatrix(treepred, as.factor(train.df$RESPONSE))
```

Correct .79 of the time on the training data


validation data testing

```{r}
treepred_2 <- predict(pruned_tr, newdata = valid.df, type = "class")
confusionMatrix(treepred_2, as.factor(valid.df$RESPONSE))
```

.73 on the validation dataset


```{r}
prp(pruned_tr, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10)
```

11 splits total


# Logistic Regression

```{r}
# make columns factor variables
germancredit$CHK_ACCT <- factor(germancredit$CHK_ACCT, levels = c(0, 1, 2, 3), labels = c("< 0", "0-200", ">200", "No checking account"))

germancredit$HISTORY <- factor(germancredit$HISTORY, levels = c(0, 1, 2, 3, 4), labels = c("no credits", "all credits paid", "credits paid until now", "delay in paying in past", "critical account"))

germancredit$SAV_ACCT <- factor(germancredit$SAV_ACCT, levels = c(0, 1, 2, 3, 4), labels = c("< 100", "100-500", "500-1000",">1000", "no savings"))

germancredit$EMPLOYMENT <- factor(germancredit$EMPLOYMENT, levels = c(0, 1, 2, 3, 4), labels = c("unemployed", "< 1", "1-4","4-7", ">7"))

germancredit$PRESENT_RESIDENT <- factor(germancredit$PRESENT_RESIDENT, levels = c(0, 1, 2, 3, 4), labels = c("0", "1", "2","3", ">4"))

germancredit$JOB <- factor(germancredit$JOB, levels = c(0, 1, 2, 3), labels = c("unemployed", "unskilled", "skilled/","management/self-employed"))

germancredit$CO.APPLICANT <- factor(germancredit$CO.APPLICANT, levels = c(0, 1), labels = c("No", "Yes"))

germancredit$GUARANTOR <- factor(germancredit$GUARANTOR, levels = c(0, 1), labels = c("No", "Yes"))

germancredit$OTHER_INSTALL <- factor(germancredit$OTHER_INSTALL, levels = c(0, 1), labels = c("No", "Yes"))

germancredit$FOREIGN <- factor(germancredit$FOREIGN, levels = c(0, 1), labels = c("No", "Yes"))

germancredit$TELEPHONE <- factor(germancredit$TELEPHONE, levels = c(0, 1), labels = c("No", "Yes"))
```

```{r}
germancredit <- germancredit[, -c(1, 5:10, 15:17, 21:22, 25:26)] # delete observation id column and merged binary variables
```


Forward Stepwise AIC For Variable Selection
```{r}
fm.0 <- logit <- glm(RESPONSE ~ 1, data = train.df, family = "binomial") 
fm.f <- step(fm.0, 
             scope = list(upper = ~ CHK_ACCT + DURATION + HISTORY + AMOUNT + SAV_ACCT + EMPLOYMENT + INSTALL_RATE + CO.APPLICANT + GUARANTOR + PRESENT_RESIDENT + AGE + OTHER_INSTALL + NUM_CREDITS + JOB + NUM_DEPENDENTS + TELEPHONE + FOREIGN + purpose_credit + marital_status + ownership_status + realestate,
                          lower = ~ 1),
             direction = "forward")
```

```{r}
fm.0 <- logit <- glm(RESPONSE ~ CHK_ACCT + DURATION + HISTORY + AMOUNT + SAV_ACCT + EMPLOYMENT + INSTALL_RATE + CO.APPLICANT + GUARANTOR + PRESENT_RESIDENT + AGE + OTHER_INSTALL + JOB + NUM_DEPENDENTS + TELEPHONE + FOREIGN + purpose_credit + marital_status + ownership_status + realestate, data = train.df, family = "binomial") 
fm.f <- step(fm.0, 
             scope = list(upper = ~ .,
                          lower = ~ CHK_ACCT),
             direction = "backward")
```

The forwards and backwards AIC result in slightly differet best models, I will use the one create by the backwards AIC because it has less variables, so it is more compact model.

```{r}
logit <- glm(RESPONSE ~ CHK_ACCT + DURATION + HISTORY + SAV_ACCT + EMPLOYMENT + INSTALL_RATE + GUARANTOR + PRESENT_RESIDENT + AGE + OTHER_INSTALL + FOREIGN + purpose_credit + marital_status + realestate, data = train.df, family = binomial(link = 'logit'))

summary(logit)
```

```{r}
pred.train <- predict(logit, newdata = train.df, type = "response")
actual.train <- train.df$RESPONSE
pred.val <- predict(logit, newdata = valid.df, type = "response")
actual.val <- valid.df$RESPONSE
```

```{r}
predClass.train <- ifelse(pred.train >0.5,1,0)

confusionMatrix(factor(actual.train), factor(predClass.train), positive = '1')
```
Using a 0.5 cutoff, the training classification rate is 0.79

```{r}
predClass.val <- ifelse(pred.val >0.5,1,0)

confusionMatrix(factor(actual.val), factor(predClass.val), positive = '1')
```

The validation classification rate is 0.70
```{r}
(FP_rate <- 29/(29+49))
```
The false positive rate is 0.27
